---
title: "Aides Drug"
author: "leia !"
date: "fall 2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.
 
```{r}
library(dplyr)
druggies = filter(triUmp, triUmp$DrugUser=='yes' )
```
 
 $H_0$ the drug will mot be effective in preventing aids in drug users 
 
 $H_A$ the drug will not be effective in preventun
 
# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->


the response variable is categorical infected based on the treatment of the new aids drug and explanatory variable is categorical treatment . therefore we utilize the numerical~cat
# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results
```{r}
barchartGC(~infected + treatment,data=druggies)
barchartGC(~infected + treatment,data=druggies, type="percent")

```

this bargraph shows that the people who got aids the ones that received the treatment vaccine was more effective than the ones that took the placebo and for the ones that took the treatment and had no aids was not effective whats so ever. 

### Graphical Descriptive Results


<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=druggies)
rowPerc(table1)
colPerc(table1)
```

the ones who where infected with aids is higher with the ones who got the placebo and the ones who got the vaccine are 3x less likely to develoop aids against those who got the placebo.  

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

the p value  0.00000001242 therefore we reject the null hypothesis 

```{r}
druggies$infected = factor(druggies$infected, c('none','aids'))
fisher.test(druggies$treatment,druggies$infected)
```

P-value 0.0000000423 the vaccine is very effective. the odds ratio is 2.81. you are almost 3 times more likely to be protected from getting aids if you get the vaccine rather then the placebo. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
 
 Is the vaccine effective in preventing aids. we did a chiqtest resulting in a p value less than 0.05. therefore we reject the null hypothesis. the vaccine is effective for preventing aids. if you take  the vaccine you are 2.8x more likely to not get aids than someone who did not take the vaccine. 
 